A survey of HTTP caching options on the open Semantic Web

### 2015-01-19 sanity/paper1.tex
move more explanations to intro
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1003 43ef1190-90fd-4dd7-a769-d0121f02b80e

RFC7234 defines two important headers: \httph{Expires}, whose value is
a date and time of when the response is considered stale, and
therefore should not be used; and \httph{Cache-Control}, which allows
detailed control of the cache, including a \httph{max-age} field,
which gives the time in seconds for how long the the response may be
used from the time of the request. \httph{max-age} takes precedence
over \httph{Expires}. In the following \emph{freshness lifetime} is
understood as the number of seconds that the response may be used
without contacting the origin server. Ideally, the calculation of the
freshness lifetime should be based of the above, but it can also be
based on heuristics, see Section~\ref{sec:simplefresh}.

RFC7232, on the other hand, defines a protocol for asking the server
if the cached response is still fresh using conditional requests. This
doesn't burden the content provider with the task of estimating the
time to live beforehand, but then, it must be able to answer if the
resource has changed less expensively than serving the entire
response. Either of two headers must be set by the server to achieve
this: \httph{ETag}, which sets an opaque identifier for the response,
or \httph{Last-Modified} which gives the time and date of the last
modification of the resource. Clients may use these values to retrieve
the updated response only if the resource has changed. Other relevant
headers are described in Table~\ref{tab:headers}
in Appendix~\ref{app:fetcher} but they were not all used in this
analysis.\todo{Structured walkthrough}\todo{Impact}

@@

The two most important headers to control caching are \httph{Expires}
and \httph{Cache-Control}\todo{move to intro}. As shown in Table~\ref{tab:headers}, there
are also others. We found \httph{Pragma} (archaic HTTP 1.0 header) in
287 responses, but except for two hosts, where they were superfluous,
they were only used to prohibit caching. \httph{Surrogates} were not
observed.

@@

Of these resources, 405 returned valid headers, but 114 did so to
\emph{prohibit} caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching.

Caching may also be prohibited by setting a non-positive freshness
lifetime. Common practice is to set the \httph{Expires} header to a
date far back in time, which has also been done in large parts of the
present dataset. For the following, we have set any non-positive
lifetime to 0, as that is semantically equivalent.

@@

We next consider the simple heuristic freshness lifetime. This is
implemented as suggested in RFC7234, Section~4.2.2. This is based on a
fraction of the time lapsed between the current time and the
modification time given in the \httph{Last-Modified} header. This
approach therefore still requires the Web server to be cooperative to
be successful, but commonly, Web servers can track this, for example
if RDF is served from a file system, the file modification time is used.

@@

\rdfterm{dct:modified} occurs in 2687 triples, but 2487 of these does
not have the Request-URI of the information resource as their subject,
i.e. it gives the last modification time of some subgraph of the
returned RDF. Nevertheless, given its prevalence, it is highly likely
that the presence of \rdfterm{dct:modified} will be useful in
determining a heuristic freshness lifetime, as the latest date may be
used.\todo{Actually calculate lifetimes based on this and compare}

>>>>

RFC7234 defines two important headers: \httph{Expires}, whose value is
a date and time of when the response is considered stale, and
therefore should not be used; and \httph{Cache-Control}, which allows
detailed control of the cache, including a \httph{max-age} field,
which gives the time in seconds for how long the the response may be
used from the time of the request. \httph{max-age} takes precedence
over \httph{Expires}. In the following \emph{freshness lifetime} is
understood as the number of seconds that the response may be used
without contacting the origin server. Ideally, the calculation of the
freshness lifetime should be based of the above, we therefore shall
refer to this as ``standards-compliant caching''. However it can also
be based on heuristics, and RFC7234, Section~4.2.2 provides some loose
constraints for such practice, as well as a suggestion for a useful
heuristic. This heuristic is based on a fraction of the time lapsed
between the current time and the modification time given in the
\httph{Last-Modified} header. This approach therefore still requires
the Web server to be cooperative to be successful, but commonly, Web
servers can track this, for example if RDF is served from a file
system, the file modification time is used.

RFC7232, on the other hand, defines a protocol for asking the server
if the cached response is still fresh using conditional requests. This
doesn't burden the content provider with the task of estimating the
time to live beforehand, but then, it must be able to answer if the
resource has changed less expensively than serving the entire
response. Either of two headers must be set by the server to achieve
this: \httph{ETag}, which sets an opaque identifier for the response,
or \httph{Last-Modified} which gives the time and date of the last
modification of the resource. Clients may use these values to retrieve
the updated response only if the resource has changed. Other relevant
headers are described in Table~\ref{tab:headers}
in Appendix~\ref{app:fetcher} but they were not all used in this
analysis.

RFC7234 provides detailed control of caching, and caching may also be
prohibited by the server, either by setting a non-positive freshness
lifetime or explicitly using a \httph{no-store} control field.

@@

As shown in Table~\ref{tab:headers}, there are also other caching
headers than those mentioned in the introduction. We found
\httph{Pragma} (archaic HTTP 1.0 header) in 287 responses, but except
for two hosts, where they were superfluous, they were only used to
prohibit caching. \httph{Surrogates} were not observed.

@@

Of these resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching.

Common practice when prohibiting caching is to set the \httph{Expires}
header to a date far back in time, which has also been done in large
parts of the present dataset. For the following, we have set any
non-positive lifetime to 0, as that is semantically equivalent.

@@

We next consider the simple heuristic freshness lifetime as suggested
RFC7234, Section~4.2.2 and detailed in the introduction.

@@

\rdfterm{dct:modified} occurs in 2687 triples, but 2487 of these does
not have the Request-URI of the information resource as their subject,
i.e. it gives the last modification time of some subgraph of the
returned RDF. Nevertheless, given its prevalence, it is highly likely
that the presence of \rdfterm{dct:modified} will be useful in
determining a heuristic freshness lifetime, as the latest date may be
used.
